---
title: "Fig4 and Fig5_14 DAG"
output: html_document
---

############################################################################################################
# Impact of inoculation practices on microbiota assembly and community stability in a fabricated ecosystem #
############################################################################################################

#clears current global environment if necessary
```{r}
rm(list = ls())
```


# Load all packages needed
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(data.table)
library(ggplot2)
library(MASS)
library(plyr)
library(multcompView)
library(ggpubr)
library(DESeq2)
library(rcompanion)
library(patchwork)
library(dplyr)
library(tidyr)
library(patchwork)
```


# Load and select 14 DAG data
```{r}
## load 16S read count table
counttab<-fread("SynCount.csv")
counttab2<-counttab[,2:ncol(counttab)]
rownames(counttab2)<-counttab$Species

## load biodiversity table
groupdiversity<-fread("Sample_Info.csv")
groupinfo2<-groupdiversity
groupinfo2<-as.data.frame(groupinfo2)
rownames(groupinfo2)<-groupinfo2$ComboInd
## Check whether count table matches group info
all(colnames(counttab2) == rownames(groupinfo2))


# select 14 DAG samples
groupinfo3 <-subset(groupinfo2, 
                    Inoculated == "Yes" & 
                      Harvest_Day =="14" &
                      Plant == "Brachypodium")
rownames(groupinfo3) <-groupinfo3$ComboInd


## this block subset the read counts from the selected groups
checkvals<-(groupinfo3$ComboInd1)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset<-append(subset,rownames(groupinfo3)[cs])}
selcols <- which(colnames(counttab2) %in% groupinfo3$ComboInd1)
counttab3<-counttab2[,..selcols] #[,rownames(groupinfo2)]
rownames(counttab3)<-rownames(counttab2)
## Check whether count table matches group info
all(colnames(counttab3) == rownames(groupinfo3))
```

# Calculate Shannin, Simpson, Richness, and Eveness
```{r}
selectedgrp = groupinfo3$G4

alpha <- with(groupinfo3, tapply(specnumber(t(counttab3)), selectedgrp, mean))
gamma <- with(groupinfo3, specnumber(t(counttab3), selectedgrp))
beta<-gamma/alpha-1

shannon<-diversity(t(counttab3))
simps<-diversity(t(counttab3),"simpson")
invsimps<-diversity(t(counttab3),"inv")
rich<-specnumber(t(counttab3))
even<-shannon/log(rich)

groupinfo3$Shannon<-shannon
groupinfo3$Simpson<-simps
groupinfo3$InvSimpson<-invsimps
groupinfo3$Richness<-rich
groupinfo3$Evenness<-even
groupinfo3 <- groupinfo3[c("Sample_Type", "Inoc_Day", "Shannon", "Simpson", "InvSimpson", "Richness", "Evenness")]
```

#set the organ appealing group in the plot
```{r}
groupinfo3$Sample_Type <- gsub("unattached sand", "sand", groupinfo3$Sample_Type)
groupinfo3$Sample_Type <- gsub("Attached sand", "rhizosphere", groupinfo3$Sample_Type)
groupinfo3$Sample_Type <- gsub("Root", "root", groupinfo3$Sample_Type)

groupinfo3$G4 <- paste(groupinfo3$Sample_Type, "_", groupinfo3$Inoc_Day)
groupinfo3$G4 <- factor(groupinfo3$G4, levels = c("sand _ 0",
                                                  "sand _ 3",
                                                  "sand _ 8",
                                                  "rhizosphere _ 0",
                                                  "rhizosphere _ 3",
                                                  "rhizosphere _ 8",
                                                  "root _ 0",
                                                  "root _ 3",
                                                  "root _ 8"))


## define groups from selected data
selectedgrp=groupinfo3$G4
```


# Fig4A, Calculate and plot alpha diversity
```{r}
## set  x, y value for ANOVA & multiple comparison
selected.Shannon <- data.frame(groupinfo3, lev=selectedgrp, y=groupinfo3$Shannon)

## this block perform ANOVA & Multiple comparison for plotting
################################################
a <- aov(y~lev, data=selected.Shannon) # perform ANOVA
tHSD <- TukeyHSD(a, ordered = FALSE, conf.level = 0.95) #perform post-hoc

generate_label_df <- function(HSD, flev){
  # Extract labels and factor levels from Tukey post-hoc 
  Tukey.levels <- HSD[[flev]][,4]
  Tukey.labels <- multcompLetters(Tukey.levels)['Letters']
  plot.labels <- names(Tukey.labels[['Letters']])
  
  # Get highest quantile for Tukey's 5 number summary and add a bit of space to buffer between    
  # upper quantile and label placement
  boxplot.df <- ddply(selected.Shannon, flev, function (x) max(fivenum(x$y)) + 0.4)
  
  # Create a data frame out of the factor levels and Tukey's homogenous group letters
  plot.levels <- data.frame(plot.labels, labels = Tukey.labels[['Letters']],
                            stringsAsFactors = FALSE)
  
  # Merge it with the labels
  labels.df <- merge(plot.levels, boxplot.df, by.x = 'plot.labels', by.y = flev, sort = FALSE)
  
  return(labels.df)
}
################################################

## This block Calculate the mean and SD of each condition/treatment pair for plotting the bar plot
################################################
agg=aggregate(groupinfo3$Shannon~selectedgrp, data=selected.Shannon, FUN="mean") #mean
agg$sd=aggregate(groupinfo3$Shannon~selectedgrp, data=selected.Shannon, FUN="sd")$value #add the SD 
library(ggplot2)
library(ggpubr)
dodge <- position_dodge(width=0.9) 
limits <- aes(ymax=groupinfo3$Shannon+sd, ymin=groupinfo3$Shannon-sd) #Set up the error bars
################################################
options(repr.plot.width = 15, repr.plot.height = 10)

## Figure 4A
Fig4A <- ggplot(groupinfo3, aes(x=selectedgrp, y=Shannon)) + 
  geom_boxplot()+
  geom_point() +
  ylim(0,2.5)+
  
  #add the ANOVA results to the plot
  #stat_compare_means(method = "anova", size=6, label.x = 1.5, label.y = 2.3) +

  #add the Post hoc results to the plot, the letter shall appear at position mean+SD+0.5
  geom_text(data = generate_label_df(tHSD, 'lev'), 
            aes(x = plot.labels, y = V1, label = labels), size=6) +
  
  #labs(title="14 DAG plant Shannon") +
  labs(y =expression("Shannon diversity index")) +
  scale_x_discrete(labels=c('0', '3', '8', '0', '3', '8', '0', '3', '8')) +
  xlab(expression(atop("_____________________          _____________________         _____________________", 
       paste("sand                                    rhizosphere                                     root")))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
Fig4A

ggsave(filename = "Fig4A.tiff", width = 6.5, height = 3, units = "in", device='tiff', dpi=300)
```


# Fig4B, Calculate and plot Beta Diversity
```{r}
#### PCoA (Bray-Cirtis dissimilarity) calculation
braydist<-vegdist(t(counttab3),method = "bray")
brayfit<-cmdscale(braydist,eig=TRUE, k=2)
eig <- brayfit$eig

brayxy<-as.data.frame(brayfit$points)
colnames(brayxy)<-c('X','Y')
points <-as.data.frame(cbind(brayxy, groupinfo3))


#### A colorblind-friendly palette for ggplot ####
# The palette with grey:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# To use for fills, add scale_fill_manual(values=cbPalette)
# To use for line and point colors, add scale_colour_manual(values=cbPalette)


#### Fig4B
options(repr.plot.width = 15, repr.plot.height = 10)

Fig4B <- ggplot(points, aes(x=X, y=Y,color=Sample_Type, 
                   shape=Inoc_Day)) +
  labs(title="PCoA (Bray-Curtis dissimilarity)") +
  labs(x=paste("PCoA 1 (", format(100 * eig[1] / sum(eig), digits=4), "%)", sep=""),
       y=paste("PCoA 2 (", format(100 * eig[2] / sum(eig), digits=4), "%)", sep="")) +
  geom_point(size=3) +
  scale_shape_manual(values = c(16, 17, 15, 3)) +
  scale_colour_manual(values=cbbPalette) +
  labs(shape="Inoculation Day (DAG)", col="Sample type") +
  theme_bw()+ theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))

Fig4B
ggsave(filename = "Fig4B.tiff", width = 6.5, height = 3, units = "in", device='tiff', dpi=300)
##################################################


# perform PERMANOVA for all 14 DAG samples
df_t_counttab3 <- as.data.frame(t(counttab3), rownames = T)
adonis2(df_t_counttab3  ~ groupinfo3$G4, data = groupinfo3, permutations = 9999, method = "bray")

adosin2_sample_type <- adonis2(df_t_counttab3  ~ groupinfo3$Sample_Type, data=groupinfo3, permutations = 9999, method = "bray")
adosin2_inoc_day <-adonis2(df_t_counttab3  ~ groupinfo3$Inoc_Day, data=groupinfo3, permutations = 9999, method = "bray")
adosin2_sample_type__inoc_day <-adonis2(df_t_counttab3  ~ groupinfo3$Sample_Type*groupinfo3$Inoc_Day, data=groupinfo3, permutations = 10000, method = "bray")

adosin2_sample_type 
adosin2_inoc_day 
adosin2_sample_type__inoc_day 

```


#Full Fig4
```{r}
options(repr.plot.width = 15, repr.plot.height = 10)

## Figure 4A
Fig4A <- ggplot(groupinfo3, aes(x=selectedgrp, y=Shannon)) + 
  geom_boxplot()+
  geom_point() +
  ylim(0,2.5)+
  
  #add the ANOVA results to the plot
  #stat_compare_means(method = "anova", size=6, label.x = 1.5, label.y = 2.7) +

  #add the Post hoc results to the plot, the letter shall appear at position mean+SD+0.5
  geom_text(data = generate_label_df(tHSD, 'lev'), 
            aes(x = plot.labels, y = V1, label = labels), size=6) +
  
  #labs(title="14 DAG plant Shannon") +
  labs(y =expression("Shannon diversity index")) +
  scale_x_discrete(labels=c('0', '3', '8', '0', '3', '8', '0', '3', '8')) +
  xlab(expression(atop("__________    __________    __________", 
       paste("sand            rhizosphere           root")))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))


Fig4 <- Fig4A/Fig4B 
Fig4 + plot_annotation(tag_levels = 'A') + plot_layout(heights = c(3, 3))

ggsave(filename = "Fig4.tiff", width = 6, height = 6, units = "in", device='tiff', dpi=300)
```


# PERMANOVA on soil, rhizosphere, and root
```{r}
#prepare the sample ID
groupinfo3$ComboInd <- rownames(groupinfo3)

# PERMANOVA on sand samples
## select 14 DAG_sand samples
sandinfo <-subset(groupinfo3, Sample_Type == "sand", drop = FALSE)
rownames(sandinfo) <-sandinfo$ComboInd
## subset the read counts from counttab3
checkvals<-(sandinfo$Sample_Type)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset<-append(subset,rownames(sandinfo)[cs])}
selcols <- which(colnames(counttab3) %in% sandinfo$ComboInd)   ###### this doesn't work
sandcounttab <- counttab3[,..selcols]
rownames(sandcounttab) <- rownames(sandcounttab)

## Check whether count table matches group info
all(colnames(sandcounttab) == rownames(sandinfo))

df_t_sandcounttab <- as.data.frame(t(sandcounttab), rownames = T) #transverse counttab3 and assigned it with species name
adosin2_sand <- adonis2(df_t_sandcounttab  ~ sandinfo$Inoc_Day, data=sandinfo, permutations = 9999, method = "bray")


# PERMANOVA on rhizosphere samples
## select 14 DAG_rhizosphere samples
rhizosphereinfo <-subset(groupinfo3, Sample_Type == "rhizosphere", drop = FALSE)
rownames(rhizosphereinfo) <-rhizosphereinfo$ComboInd
## subset the read counts from counttab3
checkvals<-(rhizosphereinfo$Sample_Type)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset<-append(subset,rownames(rhizosphereinfo)[cs])}
selcols <- which(colnames(counttab3) %in% rhizosphereinfo$ComboInd)   ###### this doesn't work
rhizospherecounttab <- counttab3[,..selcols]
rownames(rhizospherecounttab) <- rownames(rhizospherecounttab)

## Check whether count table matches group info
all(colnames(rhizospherecounttab) == rownames(rhizosphereinfo))

df_t_rhizospherecounttab <- as.data.frame(t(rhizospherecounttab), rownames = T) #transverse counttab3 and assigned it with species name
adosin2_rhizosphere <- adonis2(df_t_rhizospherecounttab  ~ rhizosphereinfo$Inoc_Day, data=rhizosphereinfo, permutations = 9999, method = "bray")


# PERMANOVA on root samples
## select 14 DAG_root samples
rootinfo <-subset(groupinfo3, Sample_Type == "root", drop = FALSE)
rownames(rootinfo) <-rootinfo$ComboInd
## subset the read counts from counttab3
checkvals<-(rootinfo$Sample_Type)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset<-append(subset,rownames(rootinfo)[cs])}
selcols <- which(colnames(counttab3) %in% rootinfo$ComboInd)   ###### this doesn't work
rootcounttab <- counttab3[,..selcols]
rownames(rootcounttab) <- rownames(rootcounttab)

## Check whether count table matches group info
all(colnames(rootcounttab) == rownames(rootinfo))

df_t_rootcounttab <- as.data.frame(t(rootcounttab), rownames = T) #transverse counttab3 and assigned it with species name
adosin2_root <- adonis2(df_t_rootcounttab  ~ rootinfo$Inoc_Day, data=rootinfo, permutations = 9999, method = "bray")


adosin2_sand
adosin2_rhizosphere
adosin2_root
```


# Fig5, DESeq on rhizosphere samples 
```{r}
# select rhizosphere data
groupinfo3$ComboInd <- rownames(groupinfo3)
groupinfo4 <- subset(groupinfo3, Sample_Type == "rhizosphere")

## this block subset the read counts from the selected groups
checkvals<-(groupinfo4$ComboInd)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset <- append(subset, rownames(groupinfo4)[cs])}
selcols <- which(colnames(counttab3) %in% groupinfo4$ComboInd)
counttab4 <- counttab3[,..selcols]
rownames(counttab4) <- rownames(counttab3)

## Check whether count table matches group info
all(colnames(counttab4) == rownames(groupinfo4))

#---------------------------------------------
# Calculate DESeq2 on 14 DAG rhizosphere data
## separate the Species name
species <- counttab[["Species"]]
species <- as.data.frame(species)
species <- separate(species, species, into = c("SynCom","bacteria", "sp", "strain"),sep = "_", remove = FALSE, extra = "merge")
species$bacname <- paste(species$bacteria, species$strain)

## give rownames to groupinfo4
row.names(groupinfo4) <- groupinfo4$ComboInd

## set `Bd21-3` as reference level
groupinfo4$Inoc_Day <- as.factor(groupinfo4$Inoc_Day)
groupinfo4$Inoc_Day <- relevel(groupinfo4$Inoc_Day, ref = "0")

## calculate DESeq2
mat_14DAG <- as.matrix(counttab4, rownames = species$bacname)
dds_14DAG <- DESeqDataSetFromMatrix(countData = mat_14DAG, colData = groupinfo4, design = ~Inoc_Day)
dds_14DAG$Inoc_Day <- factor(dds_14DAG$Inoc_Day)
dds_14DAG <- DESeq(dds_14DAG)

# Pull out DESeq2 data
alpha = 0.05

res1 = results(dds_14DAG, contrast=c("Inoc_Day", "3", "0"), alpha=alpha)
res1 = res1[order(res1$padj, na.last=NA), ]
res1$Inoc_Day <- list(rep("1"))
res_sig1 = res1[(res1$padj < alpha), ]
res_sig1

res2 = results(dds_14DAG, contrast=c("Inoc_Day", "8", "0"), alpha=alpha)
res2 = res2[order(res2$padj, na.last=NA), ]
res2$Inoc_Day <- list(rep("2"))
res_sig2 = res2[(res2$padj < alpha), ]
res_sig2

#---------------------------------------------
## set `Bd21-3` as reference level
groupinfo4$Inoc_Day <- as.factor(groupinfo4$Inoc_Day)
groupinfo4$Inoc_Day <- relevel(groupinfo4$Inoc_Day, ref = "3")

## calculate DESeq2
mat_14DAG2 <- as.matrix(counttab4, rownames = species$bacname)
dds_14DAG2 <- DESeqDataSetFromMatrix(countData = mat_14DAG2, colData = groupinfo4, design = ~Inoc_Day)
dds_14DAG2$Inoc_Day <- factor(dds_14DAG2$Inoc_Day)
dds_14DAG2 <- DESeq(dds_14DAG2)


res3 = results(dds_14DAG2, contrast=c("Inoc_Day", "8", "3"), alpha=alpha)
res3 = res3[order(res3$padj, na.last=NA), ]
res3$Inoc_Day <- list(rep("3"))
res_sig3 = res3[(res3$padj < alpha), ]
res_sig3

#---------------------------------------------
res1 <- as.data.frame(res1)
res2 <- as.data.frame(res2)
res3 <- as.data.frame(res3)

#---------------------------------------------
#### Fig5
Fig5A <- ggplot(res1, aes(x=rownames(res1), y=log2FoldChange))+ 
  geom_point(shape=21, stroke=1, aes(size=baseMean, fill = cut(padj, c(-Inf, 0.05, Inf)))) + 
  scale_fill_manual(name = "padj",values = c("(-Inf,0.05]" = "black","(0.05, Inf]" = "white")) +
  geom_hline(yintercept=0)+
  scale_size(guide=FALSE)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  theme(axis.text.x = element_blank(),
        legend.text = element_text(face="italic"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()
  ) +
  labs(title=bquote(bold("A")~"3 DAG vs 0 DAG"))

Fig5B <- ggplot(res2, aes(x=rownames(res2), y=log2FoldChange))+ 
  geom_point(shape=21, stroke=1, aes(size=baseMean, fill = cut(padj, c(-Inf, 0.05, Inf)))) + 
  scale_fill_manual(name = "padj",values = c("(-Inf,0.05]" = "black","(0.05, Inf]" = "white")) +
  geom_hline(yintercept=0)+
  scale_size(guide=FALSE)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(log[2]~fold~change))+
  theme(axis.text.x = element_blank(),
        legend.text = element_text(face="italic"),
        axis.title.x = element_blank()) +
  labs(title=bquote(bold("B")~"8 DAG vs 0 DAG"))

Fig5C <- ggplot(res3, aes(x=rownames(res3), y=log2FoldChange))+ 
  geom_point(shape=21, stroke=1, aes(size=baseMean, fill = cut(padj, c(-Inf, 0.05, Inf)))) + 
  scale_fill_manual(name = "padj",values = c("(-Inf,0.05]" = "black","(0.05, Inf]" = "white")) +
  geom_hline(yintercept=0)+
  scale_size(guide=FALSE)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  theme(axis.text.x = element_text(face="italic"),
        legend.text = element_text(face="italic"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()
  ) +
  labs(title=bquote(bold("C")~"8 DAG vs 3 DAG"))


Fig5 <- Fig5A/Fig5B/Fig5C
Fig5
ggsave(filename = "Fig5.tiff", width = 3.25, height = 6, units = "in", device='tiff', dpi=300)
```
