---
title: "Fig6, Fig7, Fig8_21DPI"
output: html_document
---

############################################################################################################
# Impact of inoculation practices on microbiota assembly and community stability in a fabricated ecosystem #
############################################################################################################

#clears current global environment if necessary
```{r}
rm(list = ls())
```


# Load all packages needed
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(data.table)
library(ggplot2)
library(MASS)
library(plyr)
library(multcompView)
library(ggpubr)
library(DESeq2)
library(rcompanion)
library(patchwork)
library(dplyr)
library(tidyr)
library(readxl)
library(tidyverse)
library(Hmisc)
```


# Load and select 21DPI data
```{r}
## load 16S read count table
counttab<-fread("SynCount.csv")
counttab2<-counttab[,2:ncol(counttab)]
rownames(counttab2)<-counttab$Species

## load biodiversity table
groupdiversity<-fread("Sample_Info.csv")
groupinfo2<-groupdiversity
groupinfo2<-as.data.frame(groupinfo2)
rownames(groupinfo2)<-groupinfo2$ComboInd
## Check whether count table matches group info
all(colnames(counttab2) == rownames(groupinfo2))

# select the 21 DPI data
groupinfo3 <-subset(groupinfo2, 
                    Inoculated == "Yes" & #"|" means "or"
                    Inoc_Assay_Length =="21" &
                    Inoc_Day !="8S" &
                    Plant == "Brachypodium" 
                    )
rownames(groupinfo3) <-groupinfo3$ComboInd
groupinfo3


## this block subset the read counts from the selected groups
checkvals<-(groupinfo3$ComboInd1)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset<-append(subset,rownames(groupinfo3)[cs])}
selcols <- which(colnames(counttab2) %in% groupinfo3$ComboInd1)
counttab3<-counttab2[,..selcols] #[,rownames(groupinfo2)]
rownames(counttab3)<-rownames(counttab2)
## Check whether count table matches group info
all(colnames(counttab3) == rownames(groupinfo3))

#set the organ appealing group in the plot
groupinfo3 <- groupinfo3[c("Sample_Type", "Inoc_Day", "Inoc_Assay_Length")]
groupinfo3$Sample_Type <- gsub("unattached sand", "sand", groupinfo3$Sample_Type)
groupinfo3$Sample_Type <- gsub("Attached sand", "rhizosphere", groupinfo3$Sample_Type)
groupinfo3$Sample_Type <- gsub("Root", "root", groupinfo3$Sample_Type)

groupinfo3$G4 <- paste(groupinfo3$Sample_Type, "_", groupinfo3$Inoc_Day)
groupinfo3$G4 <- factor(groupinfo3$G4, levels = c("sand _ 0",
                                                  "sand _ 3",
                                                  "sand _ 8",
                                                  "rhizosphere _ 0",
                                                  "rhizosphere _ 3",
                                                  "rhizosphere _ 8",
                                                  "root _ 0",
                                                  "root _ 3",
                                                  "root _ 8"))


## define groups from selected data
selectedgrp=groupinfo3$G4
```


# Calculate Shannin, Simpson, Richness, and Eveness
```{r}
selectedgrp = groupinfo3$G4

alpha <- with(groupinfo3, tapply(specnumber(t(counttab3)), selectedgrp, mean))
gamma <- with(groupinfo3, specnumber(t(counttab3), selectedgrp))
beta<-gamma/alpha-1

shannon<-diversity(t(counttab3))
simps<-diversity(t(counttab3),"simpson")
invsimps<-diversity(t(counttab3),"inv")
rich<-specnumber(t(counttab3))
even<-shannon/log(rich)

groupinfo3$Shannon<-shannon
groupinfo3$Simpson<-simps
groupinfo3$InvSimpson<-invsimps
groupinfo3$Richness<-rich
groupinfo3$Evenness<-even
```


# Fig6A, Calculate and plot alpha diversity
```{r}
## set  x, y value for ANOVA & multiple comparison
selected.Shannon <- data.frame(groupinfo3, lev=selectedgrp, y=groupinfo3$Shannon)

## this block perform ANOVA & Multiple comparison for plotting
################################################
a <- aov(y~lev, data=selected.Shannon) # perform ANOVA
tHSD <- TukeyHSD(a, ordered = FALSE, conf.level = 0.95) #perform post-hoc

generate_label_df <- function(HSD, flev){
  # Extract labels and factor levels from Tukey post-hoc 
  Tukey.levels <- HSD[[flev]][,4]
  Tukey.labels <- multcompLetters(Tukey.levels)['Letters']
  plot.labels <- names(Tukey.labels[['Letters']])
  
  # Get highest quantile for Tukey's 5 number summary and add a bit of space to buffer between    
  # upper quantile and label placement
  boxplot.df <- ddply(selected.Shannon, flev, function (x) max(fivenum(x$y)) + 0.4)
  
  # Create a data frame out of the factor levels and Tukey's homogenous group letters
  plot.levels <- data.frame(plot.labels, labels = Tukey.labels[['Letters']],
                            stringsAsFactors = FALSE)
  
  # Merge it with the labels
  labels.df <- merge(plot.levels, boxplot.df, by.x = 'plot.labels', by.y = flev, sort = FALSE)
  
  return(labels.df)
}
################################################

## This block Calculate the mean and SD of each condition/treatment pair for plotting the bar plot
################################################
agg=aggregate(groupinfo3$Shannon~selectedgrp, data=selected.Shannon, FUN="mean") #mean
agg$sd=aggregate(groupinfo3$Shannon~selectedgrp, data=selected.Shannon, FUN="sd")$value #add the SD 
library(ggplot2)
library(ggpubr)
dodge <- position_dodge(width=0.9) 
limits <- aes(ymax=groupinfo3$Shannon+sd, ymin=groupinfo3$Shannon-sd) #Set up the error bars
################################################

Fig6A <- ggplot(groupinfo3, aes(x=selectedgrp, y=Shannon)) + 
  geom_boxplot()+
  geom_point() +
  ylim(0,3)+
  
  #add the ANOVA results to the plot
  stat_compare_means(method = "anova", size=6, label.x = 1.5, label.y = 2.7) +

  #add the Post hoc results to the plot, the letter shall appear at position mean+SD+0.5
  geom_text(data = generate_label_df(tHSD, 'lev'), 
            aes(x = plot.labels, y = V1, label = labels), size=6) +
  
  labs(title="21 DPI plant Shannon") +
  labs(y =expression("Shannon Diversity Index")) +
  scale_x_discrete(labels=c('0', '3', '8', '0', '3', '8', '0', '3', '8')) +
  xlab("sand                                    rhizosphere                                     root") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))

Fig6A 
```


# Fig6B, start plotting the biodiversity chart
```{r}
#### PCoA (Bray-Cirtis dissimilarity) calculation
braydist<-vegdist(t(counttab3),method = "bray")
brayfit<-cmdscale(braydist,eig=TRUE, k=2)
eig <- brayfit$eig

brayxy<-as.data.frame(brayfit$points)
colnames(brayxy)<-c('X','Y')
points <-as.data.frame(cbind(brayxy, groupinfo3))


#### A colorblind-friendly palette for ggplot ####
# The palette with grey:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# To use for fills, add scale_fill_manual(values=cbPalette)
# To use for line and point colors, add scale_colour_manual(values=cbPalette)


#### plot PCoA (Bray-Curtis) with PCoA 1 and PCoA 2 ####
options(repr.plot.width = 15, repr.plot.height = 10)

Fig6B <- ggplot(points, aes(x=X, y=Y,color=Sample_Type, 
                   shape=Inoc_Day)) +
  labs(title="PCoA (Bray-Curtis dissimilarity)") +
  labs(x=paste("PCoA 1 (", format(100 * eig[1] / sum(eig), digits=4), "%)", sep=""),
       y=paste("PCoA 2 (", format(100 * eig[2] / sum(eig), digits=4), "%)", sep="")) +
  geom_point(size=3) +
  scale_shape_manual(values = c(6, 16, 17)) +
  scale_colour_manual(values=cbbPalette) +
  labs(shape="Inoculation day (DAG)", col="Sample type") +
  theme_bw()
Fig6B
##################################################

## perform PERMANOVA
df_t_counttab3 <- as.data.frame(t(counttab3), rownames = T)
adonis2(df_t_counttab3  ~ Sample_Type*Inoc_Day, data=groupinfo3, permutations = 10000, method = "bray")
```


#Full Fig6
```{r}
options(repr.plot.width = 15, repr.plot.height = 10)

## Figure 6A
Fig6A <- ggplot(groupinfo3, aes(x=selectedgrp, y=Shannon)) + 
  geom_boxplot()+
  geom_point() +
  ylim(0,3)+
  
  #add the ANOVA results to the plot
  stat_compare_means(method = "anova", size=6, label.x = 1.5, label.y = 2.7) +

  #add the Post hoc results to the plot, the letter shall appear at position mean+SD+0.5
  geom_text(data = generate_label_df(tHSD, 'lev'), 
            aes(x = plot.labels, y = V1, label = labels), size=6) +
  
  labs(y =expression("Shannon diversity index")) +
  scale_x_discrete(labels=c('0', '3', '8', '0', '3', '8', '0', '3', '8')) +
  xlab(expression(atop("__________    __________    __________", 
       paste("sand            rhizosphere           root")))) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))


Fig6 <- Fig6A/Fig6B 
Fig6 + plot_annotation(tag_levels = 'A') + plot_layout(heights = c(3, 3))

ggsave(filename = "Fig6.tiff", width = 6, height = 6, units = "in", device='tiff', dpi=300)
```


# DESeq2 between rhizosphere vs sand (1st comparison)
```{r}
## separate the Species name
species <- counttab[["Species"]]
species <- as.data.frame(species)
species <- separate(species, species, into = c("SynCom","bacteria", "sp", "strain"),sep = "_", remove = FALSE, extra = "merge")
species$bacname <- paste(species$bacteria, species$strain)

#---------------------------------------------
# Step1: Select rhizosphere and sand data for DESeq2
groupinfo3$ComboInd <- rownames(groupinfo3)
noroot <- subset(groupinfo3, Sample_Type != "root")

## this block subset the read counts from the selected groups
checkvals<-(noroot$ComboInd)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset <- append(subset, rownames(noroot)[cs])}
selcols <- which(colnames(counttab3) %in% noroot$ComboInd)
counttab_noroot <- counttab3[,..selcols]
rownames(counttab_noroot) <- rownames(counttab3)

## Check whether count table matches group info
all(colnames(counttab_noroot) == rownames(noroot))

## give rownames to noroot
row.names(noroot) <- noroot$ComboInd

## set `sand` as reference level
noroot$Sample_Type <- as.factor(noroot$Sample_Type)
noroot$Sample_Type <- relevel(noroot$Sample_Type, ref = "sand")

## calculate DESeq2
mat_noroot <- as.matrix(counttab_noroot, rownames = species$bacname)
dds_noroot <- DESeqDataSetFromMatrix(countData = mat_noroot, colData = noroot, design = ~Sample_Type)
dds_noroot$Sample_Type <- factor(dds_noroot$Sample_Type)
dds_noroot <- DESeq(dds_noroot)
resultsNames(dds_noroot)

# Pull out DESeq2 data
alpha = 0.05
res1 = results(dds_noroot, name = "Sample_Type_rhizosphere_vs_sand", alpha=alpha)
res1_all = res1[order(res1$padj), ]
res1 = res1[order(res1$padj, na.last=NA), ]
res_sig1 = res1[(res1$padj < alpha), ]
res_sig1


#---------------------------------------------
# Step2: Select root and sand data for DESeq2
groupinfo3$ComboInd <- rownames(groupinfo3)
colData2 <- subset(groupinfo3, Sample_Type != "rhizosphere")

## this block subset the read counts from the selected groups
checkvals<-(colData2$ComboInd)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset <- append(subset, rownames(colData2)[cs])}
selcols <- which(colnames(counttab3) %in% colData2$ComboInd)
countData2 <- counttab3[,..selcols]
rownames(countData2) <- rownames(counttab3)

## Check whether count table matches group info
all(colnames(countData2) == rownames(colData2))

## give rownames to colData2
row.names(colData2) <- colData2$ComboInd

# set `sand` as reference level
colData2$Sample_Type <- as.factor(colData2$Sample_Type)
colData2$Sample_Type <- relevel(colData2$Sample_Type, ref = "sand")

## calculate DESeq2
mat2 <- as.matrix(countData2, rownames = species$bacname)
dds2 <- DESeqDataSetFromMatrix(countData = mat2, colData = colData2, design = ~Sample_Type)
dds2$Sample_Type <- factor(dds2$Sample_Type)
dds2 <- DESeq(dds2)
resultsNames(dds2)

# Pull out DESeq2 data
res2 = results(dds2, name = "Sample_Type_root_vs_sand", alpha = 0.05)
res2_all = res2[order(res2$padj), ]
res2 = res2[order(res2$padj, na.last=NA), ]
alpha = 0.05
res_sig2 = res2[(res2$padj < alpha), ]
res_sig2


#---------------------------------------------
# Step3: Select sand and sand data for DESeq2
groupinfo3$ComboInd <- rownames(groupinfo3)
nosand <- subset(groupinfo3, Sample_Type != "sand")

## this block subset the read counts from the selected groups
checkvals<-(nosand$ComboInd)
subset<-list()
countset<-1:length(checkvals)
for (cs in countset){subset <- append(subset, rownames(nosand)[cs])}
selcols <- which(colnames(counttab3) %in% nosand$ComboInd)
counttab_nosand <- counttab3[,..selcols]
rownames(counttab_nosand) <- rownames(counttab3)

## Check whether count table matches group info
all(colnames(counttab_nosand) == rownames(nosand))

## give rownames to nosand
row.names(nosand) <- nosand$ComboInd

# set `rhizosphere` as reference level
nosand$Sample_Type <- as.factor(nosand$Sample_Type)
nosand$Sample_Type <- relevel(nosand$Sample_Type, ref = "rhizosphere")

## calculate DESeq2
mat_nosand <- as.matrix(counttab_nosand, rownames = species$bacname)
dds_nosand <- DESeqDataSetFromMatrix(countData = mat_nosand, colData = nosand, design = ~Sample_Type)
dds_nosand$Sample_Type <- factor(dds_nosand$Sample_Type)
dds_nosand <- DESeq(dds_nosand)
resultsNames(dds_nosand)


# Pull out DESeq2 data
alpha = 0.05
res3 = results(dds_nosand, name = "Sample_Type_root_vs_rhizosphere", alpha=alpha)
res3_all = res3[order(res3$padj), ]
res3 = res3[order(res3$padj, na.last=NA), ]
res_sig3 = res3[(res3$padj < alpha), ]
res_sig3
```


# Fig7, Sig. diff. DESeq2 results
```{r}
# Visualizing the data
res_sig1 <- as.data.frame(res_sig1)
res_sig2 <- as.data.frame(res_sig2)
res_sig3 <- as.data.frame(res_sig3)


Fig7A <- ggplot(res_sig1, aes(x=rownames(res_sig1), y=log2FoldChange))+ 
  geom_point(shape=21, stroke=1, aes(size=baseMean, fill = cut(padj, c(-Inf, 0.05, Inf)))) + 
  scale_fill_manual(name = "padj",values = c("(-Inf,0.05]" = "black","(0.05, Inf]" = "white")) +
  geom_hline(yintercept=0)+
  scale_size(guide=FALSE)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(log[2]~fold~change))+
  theme(legend.text = element_text(face="italic"),
        axis.title.x = element_blank()) +
  labs(title=bquote("rhizosphere vs sand"))


Fig7B <- ggplot(res_sig3, aes(x=rownames(res_sig3), y=log2FoldChange))+ 
  ylim(-7, 1) +
  geom_point(shape=21, stroke=1, aes(size=baseMean, fill = cut(padj, c(-Inf, 0.05, Inf)))) + 
  scale_fill_manual(name = "padj",values = c("(-Inf,0.05]" = "black","(0.05, Inf]" = "white")) +
  geom_hline(yintercept=0)+
  scale_size(guide=FALSE)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.position="none")+
  geom_hline(yintercept=0)+
  ylab(expression(log[2]~fold~change))+
  theme(legend.text = element_text(face="italic"),
        axis.title.x = element_blank()) +
 labs(title=bquote("root vs rhizosphere"))

Fig7 <- Fig7A + Fig7B 
Fig7 + plot_annotation(tag_levels = 'A') + plot_layout(widths = c(5, 3))
ggsave(filename = "Fig7.tiff", width = 6.5, height = 3, units = "in", device='tiff', dpi=300)
```


# Fig8, Calculate_reproducibility among replicates based on Sample_Type
```{r}
# Load 16S read count table
logrelabuntab<-fread("Syn_logRelCount.csv")
logrelabuntab2<-logrelabuntab[,2:ncol(logrelabuntab)]
rownames(logrelabuntab2)<-logrelabuntab$Species
braydist<-vegdist(t(logrelabuntab2),method = "bray")
braydist2 <- data.frame(as.matrix(braydist))
points <-as.data.frame(cbind(braydist2, groupinfo2))

## Check whether the name is correctly aligned
all(rownames(points) == rownames(groupinfo2)) 

# extract the numeric data
df <- read_xlsx("SynBrayDist.xlsx")
df_element <- df[, 2:171] 
colnames(df_element) <- NULL
df_element2 <- matrix(unlist(t(df_element)), byrow = FALSE, ncol = 1) %>% as.data.frame()
df_name <- df %>% dplyr::select(G4)

# Combine data with name
df_name <- do.call("rbind", replicate(n = nrow(df_element) / nrow(df_name), df_name, simplify = FALSE))
df_name2 <- rep(df$G4, each=170)

# Delete repetitive data
df_all <- cbind(df_name, df_name2, df_element2) %>% filter(G4 == df_name2) %>% distinct(V1, .keep_all = TRUE) %>% filter(V1 != 0)
df_all <-data.frame(df_all)
colnames(df_all) <-c("G4", "group", "braydist")


# Use for loop to plot reproducibility according to sample type
## Set the organ appealing group in the plot
df_all$group <- gsub("unattached sand", "sand", df_all$group)
df_all$group <- gsub("Attached sand", "rhizosphere", df_all$group)
df_all$group <- gsub("Root", "root", df_all$group)

## Set sample type
type <- c("sand", "rhizosphere", "root")
'%notlike%' <- Negate('%like%')
mylist <- list()

for (sample_type in type) {
  df_subset <-subset(df_all, 
                     group %like% 21 & 
                     group %like% sample_type &
                     group %notlike% 'None' &
                     group %notlike% 'No' &
                     group %notlike% '8S')

## Perform ANOVA & Multiple comparison for plotting
d <- data.frame(lev=df_subset$group, y=df_subset$braydist)
a <- aov(y~lev, data=d)
tHSD <- TukeyHSD(a, ordered = FALSE, conf.level = 0.95)
  
generate_label_df <- function(HSD, flev){
    Tukey.levels <- HSD[[flev]][,4]
    Tukey.labels <- multcompLetters(Tukey.levels)['Letters']
    plot.labels <- names(Tukey.labels[['Letters']])
    boxplot.df <- ddply(d, flev, function (x) max(fivenum(x$y)) + 0.4)
    plot.levels <- data.frame(plot.labels, labels = Tukey.labels[['Letters']], stringsAsFactors = FALSE)
    labels.df <- merge(plot.levels, boxplot.df, by.x = 'plot.labels', by.y = flev, sort = FALSE)
    return(labels.df)
}

## Calculate the mean and SD of each condition/treatment pair for plotting the bar plot
agg=aggregate(y~lev, data=d, FUN="mean")
agg$sd=aggregate(y~lev, data=d, FUN="sd")$value
library(ggplot2)
dodge <- position_dodge(width=0.9) 
limits <- aes(ymax=value+sd, ymin=value-sd)
  
#set  x, y value for ANOVA & multiple comparison
d <- data.frame(lev=df_subset$group, y=df_subset$braydist)
  

## Fig8
draw_data <- ggplot(d, aes(x=lev, y=y)) + 
    geom_boxplot()+
    geom_point() +
    ylim(0, 0.55) +
    #add the ANOVA results to the plot
    stat_compare_means(method = "anova", size=3.5, label.x = 1, label.y = 0.5) +
    #add the Post hoc results to the plot, the letter shall appear at position mean+SD+0.5
    geom_text(data = generate_label_df(tHSD, 'lev'), 
              aes(x = plot.labels, y = V1-0.3, label = labels), size=6) +
    #setup texts font, size, angle in the graph
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(title = element_text(face = "bold", size = 8))+ #set title format
    theme(axis.text.x = element_text(face = "bold", color="black", size = 8),
          axis.text.y = element_text(face = "bold", color="black", size = 8),
          plot.margin = unit(c(0.2, 0.2, 0.3, 0.3), "inches")) + #set x, y axis format, and plot margin
    scale_x_discrete(labels=c("0", "3", "8")) +
    labs(title = paste(sample_type),
         x =expression(" "), 
         y =expression(bold("Euclidean distances among replicates")))
mylist[[sample_type]] <- draw_data
}

myplot <- patchwork::wrap_plots(mylist, nrow=1) + patchwork::plot_annotation(tag_levels = 'A')
myplot
ggsave(filename = "Fig8.tiff", myplot, width = 6.5, height = 3, units = "in", device='tiff', dpi=300)
```
